Leadership across cultures and contexts: an exploratory analysis

### 2020-05-12 leadership_across_cultures_contexts.Rmd
Edits to final discussion sections.

Some common leadership dimensions were also relatively invariant across the four factors we examined: group context, subsistence strategy, continental region, and leader sex. These dimensions included qualities, such as being knowledgeable or intelligent, wealthy, strategically nepotistic, and generous, and functions, such as resolving conflict and representing the group. These dimensions are therefore candidates for universal dimensions of human leadership. By "universal" we do not mean that all leaders have all these traits. Instead, we mean that there is evidence for these traits across a solid majority of cultures that do not vary much by key measures of population or group varition (i.e., the four factors). See Table \@ref(tab:leaderuniversals) and Figure \@ref(fig:varsupport).

@@

The universal dimensions also partially correspond to the service-for-prestige model [@price_evolution_2014; @price_service-for-prestige_2015] in which leaders receive special status in exchange for the services they provide followers, but that model does not emphasize the concrete material benefits seen here. The absence of *High status* and *Aggressiveness* on this list of universals corresponds to the reverse dominance hierarchy model [@boehm_egalitarian_1993], which emphasizes knowledge, intelligence, conflict resolution, generosity, and status leveling mechanisms among largely egaliarian societies.

@@

The clearest pattern to emerge from analyses of systematic variation is that, not too surprisingly, many dimensions of leadership are context-specific: aggressiveness and bravery were associated with military leadership, supernatural qualities and ritual functions were associated with religious leadership, organizing cooperation and being experience or accomplished were associated with economic-group leadership, and providing counsel and older age were associated with kin-group leadership. See Figure \@ref(fig:heatmapgroups). These results underscore recent trends in evoltuionary approaches to leadership that emphasize the context-dependency of leader emergence and evolutionary selection pressures [@smith_leadership_2020; @smith_leadership_2016; @hagen_leadership_2019].

Four leadership dimensions varied by subsistence strategy. There was more evidence that hunter-gatherer leaders lacked coercive authority compared to other leaders, which supports much theory and ethnography discussing egalitarianism and the active resistance of hierarchy among hunter-gatherers [e.g., @boehm_hierarchy_1999;@boehm_purposive_2008; @woodburn_egalitarian_1982]. Reliance on a mobile and stochastic resource base is suggested to promote strong social norms of sharing [@peterson_demand_1993; @cashdan_egalitarianism_1980] and shape both resource transfers and partner preferences [c.f., @smith_friend_2018; @smith_hunter-gatherers_2018]. There was more evidence that both hunter-gatherer and pastoralist leaders generated food surpluses, but less evidence that pastoralist leaders distributed food and other resources. Thus, systematic ethnographic evidence supports the important role of provisioning resources by influential males among foragers [@wood_prestige_2006; @gurven_give_2005; @gurven_men_2009; @alger_paternal_2020]. However, there was more evidence that leaders in communities with mixed subsistence strategies distributed resources, which lends support to males using economic success and producitivity as an influence-seeking strategy [@hawkes_showing_2002; @hawkes_family_2010].

@@

Relationships between leadership and high status, including causality and dependence, are puzzling and currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_why_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019; @garfield_evolutionary_2019]. Drawing on a comparative-evolutionary perspective, @van_vugt_dual_2019 emphasize that although leadership and high social status do often covary, leaders are not always high-status individuals, exemplified by low-ranking individuals in nonhuman species that lead group movement or younger and lower status men and women who initiate collective actions. @cheng_dominance_2019 and @cheng_why_2020 emphasize, however, that among humans high social status may most often be a necessary but insufficient condition for achieving and maintaining positions of leadership. The "low status" young adults who initiate a collective action might be high status among their peers, within what @garfield_review_2019 [p. 71] described as the "intragenerational social hierarchy."

In our view, social context is critical to these discussions [see @wiessner_power_2010]. Leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social "statuses." In many cultures, for example, communities form work-groups and appoint "chiefs-for-a-day" who organize the work but have little status or authority otherwise [e.g., @macfarlan_competitive_2012]. @garfield_evolutionary_2019 found that leader organization variables clustered separately from the leader prestige variables, suggesting they are distinct. @garfield_investigating_2019 found that some leaders in recently settled hunter-gatherers had organizing responsibilities but unexceptional levels of respect. Some older individuals, on the other hand, were highly respected but had little social influence or authority.

In our results, high status was the most common quality of leaders at both the text record  (`r statusPercTxt`%) and culture level (`r statusPercCult`%) (Figure \@ref(fig:varsupport)); it was also the second most common benefit of leadership, and the loss of status was the most frequent cost (Figure \@ref(fig:varsupportcb)).
<!-- mentioned in this sample of the ethnographic record which includes discussions of chiefs, work group leaders, shamans, and age-sets, from a culturally diverse sample including foraging groups and kingdoms.  -->
This lends support to the notion of high social status often being a necessary condition of leader emergence [@cheng_why_2020]. Nevertheless, over two thirds of our text records on leadership did not mention status, and it was not one of our candidate universal dimensions of leadership. Interestingly, evidence for high status leaders varied by continental region, with more evidence for them from South American cultures (see Figure S\@ref(fig:heatmapregion)).

@@

When ethnographic texts described female leaders exclusively, they were relatively more likely to be described as married to a high-quality spouse in a polygynous marriage, and the recipients of various mating benefits (Figure S\@ref(fig:heatmapsex)). Female leaders also had higher levels of evidence for the social, reproductive, and material success "feature" variable (Figure S\@ref(fig:heatmapfeaturesex)). Evidence for female leaders was also more likely to be found in kin groups and residential subgroups. This suite of characteristics is consistent with discussions of high status first-wives who gain and maintain increased social influence across the lifespan by leveraging their extended kin and social networks [@goodale_tiwi_1971; @brown_her_1985; @yanca_female_2004]. @garfield_investigating_2019 similarly found that there was positive assortative mating of high status leaders among elected leaders in a small-scale society with relative gender-egalitarianism. Overall, our resuls are also consistent with theoretical arguments emphasizing women as leaders of families [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019].

@@

Our systematic cross-cultural results support many aspects of influential evolutionary theories of leadership.

**Dominance and prestige**. Much of the evolutionary theoretical literature on leadership has suggested social influence is most often achieved via one of two distinct strategies: dominance or prestige [for review see @garfield_review_2019; @garfield_evolutionary_2019]. The dominance-prestige theory as formulated by @henrich_evolution_2001 significantly advanced evolutionary approaches to leadership and social hierarchy and inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @maner_2016; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019; @garfield_investigating_2019]. @van_vugt_dual_2019 [p. 1], for example, "argue that there are two leadership styles: prestige style and dominance style," and @cheng_two_2013 [p. 119] in focusing on distinctions between dominance and prestige suggests, "human social stratification is characterized by the co-occurrence of both strategies." Exploratory cluster analyses reported here, however, suggest multiple features of leader qualities and leader functions, beyond dominance and prestige. Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, in order to organize group members, implement strategic actions, and provide prosocial services to the group, all while conforming to cultural norms. Dichotomizing this complexity into two distinct social strategies, although theoretically and empirically convenient, likely fails to capture much of the ethnographic reality of human leadership dynamics.

@@

<!-- DISCUSSES SUCCESS FEATURE -->
**Leader success**. Leaders, probably as a consequence of their ability to provide social benefits and impose social costs, are also characterized by their material, social, and reproductive success (bottom cluster Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature focused on associated outcomes of leadership and social hierarchy [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013]. Several of these measures of success often co-occur. Among the Trobriands, horticulturalists in Papua New Guinea, @ol06-037 [p. 47] describes relationships between polygynous marriages, increased wealth, and widespread social contacts:

<!-- DISCUSSES COMPETENCIES FEATURE -->
The individual-level competencies feature was comprised of variables associated with dominance-based leadership, including coercive authority and aggressiveness [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013] as well as a group that included qualities of knowledgeable and intelligent and experienced and accomplished, traits that have been emphasized by the prestige theory [@henrich_evolution_2001; @cheng_two_2013] and the more general computational services theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and which are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)). Interestingly, being feared and having supernatural qualities clustered more closely with the prestige-style traits than with the dominance-style traits. There is likely a strong connection between shamanism and ritual-based influence with both prestige and dominance styles of leadership. Ritual and supernatural capacities may represent a dimension of leader qualities in which dominance and prestige are coupled forming an alternative style of leadership, as shamans often provide a necessary and valued public good, have special expertise and abilities, are feared and achieve or maintain influence via threats or use of direct and/or indirect aggression. Artistic performance is also implicated in this style of leadership and ceremonial, artistic performances by shamans are common displays which often incorporate the superhuman abilities of shamans [@singh_cultural_2017].

@@

<!-- PROSCIAL FUNCTIONS FEATURE -->
The prosocial feature of leader functions included distributing resources, providing protection, and various prosocial investments, which are functions commonly highlighted in the anthropological literature on leadership in nonindustrial societies [@sahlins_poor_1963; @johnson_organizational_1982; @service_origins_1975; @boehm_egalitarian_1993; @boehm_hierarchy_1999; @johnson_evolution_1987]. In the absence of institutionalized leadership roles, which characterize many of the relatively small-scale populations in this sample, leaders often maintain social influence via continued community support and the demands of followers can shape the functions of leaders to meet local needs.

<!-- DISCUSSES BOTH MANAGEMENT FEATURES -->
In the absence of state-institutions local leaders are often tasked with managing various economic systems and providing social services. Within the domain of leader management, there are likely to be two primary features of human leader functions: strategizing and organizing. Our results correspond very closely to what @van_vugt_cognitive_2007 define as *strategic* versus *coordinating* leadership.

@@

@@

[NEED TO INTEGRATE]
These results provide a framework for further work on evolutionarily relevent costs and benfits of leadership, and suggests leaders-follower dynamics should continue to be prioritized in evolutionary models of demography, group dynamics, social complexity and inequality, cooperation and punishment, and resource exchange [c.f., @bird_variability_2019; @ember_our_2018; @micheletti_modelling_2020; @smith_cultural_2020; @birch2017philosophy; @marlowe_more_2008; @henrich_markets_2010; @mace_co-evolution_1998; @mace_phylogenetic_2005; @mattison_evolution_2016]

@@

# Concluding remarks: Beyond the non-WEIRD 
<!-- NOT SURE OF TITLE -- WANT TO EMPHASIZE NON WEIRD DIVERSITY AND ALSO MENTION BACK TO WERID FINDINGS BRIEFLY -->

Leaders in the ethnographic record are portrayed as generally prosocial, serve to resolve conflicts and organize cooperation, and receive material, social, and reproductive benefits. There is also a "dominant" style of leadership involving physical formidability, fear, and aggression. These results conform to the characteristics of leaders that are emphasized in the literature on the evolution of leadership and provide broad cross-cultural support for several influential theories, such as the prestige-dominance model, the collective action model, the service-for-prestige model, and the computational services model [for review, see @garfield_evolutionary_2019].

One interesting pattern that has not been widely discussed in the evolutionary literature on leadership is the role of shamans and their intersection, in our results, with the dominant style of leadership. Our social costs cluster (Figure \@ref(fig:cluster)A) includes the *Supernatural* abilities variable. @winkelman1987cross found that in 47% of societies sampled from the HRAF ($n = 43$) shamans were associated with providing decision-making services and associated with military power. Providing ritual functions and military leadership are common (documented in over 50% of societies here) functions leaders serve (see Figure \@ref(fig:varsupport)). Widespread preferences for egalitarianism and social equality are likely strong drivers of leader preferences across diverse societies [@cheng_dominance_2019], but shamanism and ritual based contexts represent a social domain in which ideologies support dominance-based leadership.

We also found a link between prestige-based leadership, which is associated with deference towards accomplished individuals with expertise, special skills, and high intelligence [@cheng_dominance_2019; @henrich_evolution_2001], and charismatic leadership, which is associated with oratory skills, charismatic personality, confidence, and interpersonal skills [@grabo_charismatic_2016; @howell_role_2005]. See, Figure \@ref(fig:cluster)A, *Ability to provide social benefits* cluster. This suggests that these leadership strategies might be closely related across a broad range of societies and social contexts, or part of a broader strategy of social influence based on the provisioning of various benefits.

The strong support for material and social benefits associated with leadership (Figure \@ref(fig:varsupportcb)) and the colinearity of measures associated with material wealth, social capital, and reproductive success (Figure \@ref(fig:cluster)A), suggests processes of leadership and followership might be strong drivers of the emergence and intergenerational transmission of inequality [@bowles_emergence_2010; @mattison_evolution_2016]. Various sources of material wealth and strategic or institutionalized processes of nepotism are widespread cross-culturally, including among many societies often described as "egalitarian." These results suggest anthropologists and economists investigating the emergence and evolution of inequality, including material, social, and reproductive dimensions, should incorporate and pay close attention to group processes and social strategies associated with leadership and followership.

@@

The 60-culture Probability Sample Files of the HRAF developed by @naroll_proposed_1967 is based on culturally and geographically diverse populations which are well covered in the ethnographic record. The distribution of cultures, however, is not equal across categories of subsistence strategies; agriculturalists and horticulturalists are relatively overpresented compared to hunter-gatherers and pastoralists. The ethnographic sample of text developed in the *leadershipdata* package is also not equally balanced across all subsistence stragey categories. See Figure S\@ref(fig:subsisttree).

@@

The distribution of cultures by subsistence stragegy across conteinetla region is also variable. In this sample, most African cultures are agriculturalists and most North American cultures are hunter-gatherers, for example. See Figure S\@ref(fig:regionsubsistence).

>>>>

Some common leadership dimensions were also relatively invariant across the four factors we examined: group context, subsistence strategy, continental region, and leader sex. These dimensions included qualities, such as being knowledgeable or intelligent, wealthy, strategically nepotistic, and generous, and functions, such as resolving conflict and representing the group. These dimensions are therefore candidates for universal dimensions of human leadership. By "universal" we do not mean that all leaders have all these traits. Instead, we mean that there is evidence for these traits across a solid majority of cultures that do not vary much by key measures of population or group variation (i.e., the four factors). See Table \@ref(tab:leaderuniversals) and Figure \@ref(fig:varsupport).

@@

The universal dimensions also partially correspond to the service-for-prestige model [@price_evolution_2014; @price_service-for-prestige_2015] in which leaders receive special status in exchange for the services they provide followers, but that model does not emphasize the concrete material benefits seen here. The absence of *High status* and *Aggressiveness* on this list of universals corresponds to the reverse dominance hierarchy model [@boehm_egalitarian_1993], which emphasizes knowledge, intelligence, conflict resolution, generosity, and status leveling mechanisms among largely egalitarian societies.

@@

The clearest pattern to emerge from analyses of systematic variation is that, not too surprisingly, many dimensions of leadership are context-specific: aggressiveness and bravery were associated with military leadership, supernatural qualities and ritual functions were associated with religious leadership, organizing cooperation and being experience or accomplished were associated with economic-group leadership, and providing counsel and older age were associated with kin-group leadership. See Figure \@ref(fig:heatmapgroups). These results underscore recent trends in evolutionary approaches to leadership that emphasize the context-dependency of leader emergence and evolutionary selection pressures [@smith_leadership_2020; @smith_leadership_2016; @hagen_leadership_2019].

Four leadership dimensions varied by subsistence strategy. There was more evidence that hunter-gatherer leaders lacked coercive authority compared to other leaders, which supports much theory and ethnography discussing egalitarianism and the active resistance of hierarchy among hunter-gatherers [e.g., @boehm_hierarchy_1999;@boehm_purposive_2008; @woodburn_egalitarian_1982]. Reliance on a mobile and stochastic resource base is suggested to promote strong social norms of sharing [@peterson_demand_1993; @cashdan_egalitarianism_1980] and shape both resource transfers and partner preferences [c.f., @smith_friend_2018; @smith_hunter-gatherers_2018]. There was more evidence that both hunter-gatherer and pastoralist leaders generated food surpluses, but less evidence that pastoralist leaders distributed food and other resources. Thus, systematic ethnographic evidence supports the important role of provisioning resources by influential males among foragers [@wood_prestige_2006; @gurven_give_2005; @gurven_men_2009; @alger_paternal_2020]. However, there was more evidence that leaders in communities with mixed subsistence strategies distributed resources, which lends support to males using economic success and productivity as an influence-seeking strategy [@hawkes_showing_2002; @hawkes_family_2010].

@@

Relationships between leadership and high status, including causality and dependence, are puzzling and currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_why_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019; @garfield_evolutionary_2019]. Drawing on a comparative-evolutionary perspective, @van_vugt_dual_2019 emphasize that although leadership and high social status do often covary, leaders are not always high-status individuals, exemplified by low-ranking individuals in nonhuman species that lead group movement or younger and lower status men and women who initiate collective actions. @cheng_dominance_2019 emphasizes, however, that among humans high social status may most often be a necessary but insufficient condition for achieving and maintaining positions of leadership [see also @cheng_why_2020]. The "low status" young adults who initiate a collective action might be high status among their peers, within what @garfield_review_2019 [p. 71] described as the "intragenerational social hierarchy."

In our view, social context is critical to these discussions [see @wiessner_power_2010]. Leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social "statuses." In many cultures, for example, communities form work-groups and appoint "chiefs-for-a-day" who organize the work but have little status or authority otherwise [e.g., @macfarlan_competitive_2012]. @garfield_evolutionary_2019 found that leader organization variables clustered separately from leader prestige variables, suggesting they are distinct. @garfield_investigating_2019 found that some leaders among recently settled hunter-gatherers had organizing responsibilities but unexceptional levels of respect. Some older individuals, on the other hand, were highly respected but had little social influence or authority.

In our results, high status was the most common quality of leaders at both the text record  (`r statusPercTxt`%) and culture level (`r statusPercCult`%) (Figure \@ref(fig:varsupport)); it was also the second most common benefit of leadership and the loss of status was the most frequent cost (Figure \@ref(fig:varsupportcb)).
<!-- mentioned in this sample of the ethnographic record which includes discussions of chiefs, work group leaders, shamans, and age-sets, from a culturally diverse sample including foraging groups and kingdoms.  -->
This lends support to the notion of high social status often being a necessary condition of leader emergence [@cheng_why_2020]. Nevertheless, over two thirds of our text records on leadership did not mention status, and it was not one of our candidate universal dimensions of leadership. Interestingly, evidence for high status leaders varied by continental region, with more evidence for them from South American cultures (see Figure S\@ref(fig:heatmapregion)).

@@

When ethnographic texts described female leaders exclusively, they were relatively more likely to be described as married to a high-quality spouse in a polygynous marriage, and the recipients of various mating benefits (Figure S\@ref(fig:heatmapsex)). Female leaders also had higher levels of evidence for the social, reproductive, and material success "feature" variable (Figure S\@ref(fig:heatmapfeaturesex)). Evidence for female leaders was also more likely to be found in kin groups and residential subgroups. This suite of characteristics is consistent with discussions of high status first-wives who gain and maintain increased social influence across the lifespan by leveraging their extended kin and social networks [@goodale_tiwi_1971; @brown_her_1985; @yanca_female_2004]. @garfield_investigating_2019 similarly found that there was positive assortative mating of high status leaders among elected leaders in a small-scale society with relative gender-egalitarianism. Overall, our results are also consistent with theoretical arguments emphasizing women as leaders of families [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019].

@@

Our systematic cross-cultural results support many aspects of influential evolutionary theories of leadership, but offer new direction for the expansion and synthesis of leadership theories.

**Dominance and prestige**. Much of the evolutionary theoretical literature on leadership has suggested social influence is most often achieved via one of two distinct strategies: dominance or prestige [for review see @garfield_review_2019; @garfield_evolutionary_2019]. There is a long history of leadership styles being described as based on either force and strength (e.g., dominance) or respect and special skills (e.g., prestige). Such dual-models of leadership styles are rooted in the coupling of ethology and ethnography [see @mead_sex_1935; @sahlins_poor_1963; @lewis_leaders_1974; @tiger_dominance_1970; @bernstein_dominance_1976]. Synthesizing the ethnographic and ethological records, @kracke_force_1978 linked dominance and prestige to psychological-affect between leaders and followers in small-scale societies and @barkow_prestige_1980 suggested the evolution of prestige was linked to sexual selection and symbolic culture. The dominance-prestige theory as formulated by @henrich_evolution_2001 significantly advanced evolutionary approaches to leadership and social hierarchy and inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @maner_2016; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019; @garfield_investigating_2019]. @van_vugt_dual_2019 [p. 1], for example, "argue that there are two leadership styles: prestige style and dominance style," and @cheng_two_2013 [p. 119] in focusing on distinctions between dominance and prestige suggests, "human social stratification is characterized by the co-occurrence of both strategies." Exploratory cluster analyses reported here, however, suggest multiple features of leader qualities and leader functions, beyond dominance and prestige.

@@

<!-- DISCUSSES COMPETENCIES FEATURE -->
**Leader capacities**. The individual-level competencies feature was comprised of variables associated with dominance-based leadership, including coercive authority and aggressiveness [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013] as well as a group that included qualities of knowledgeable and intelligent and experienced and accomplished, traits that have been emphasized by the prestige theory [@henrich_evolution_2001; @cheng_two_2013] and the more general computational services theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and which are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)). Interestingly, being feared and having supernatural qualities clustered more closely with the prestige-style traits than with the dominance-style traits. There is likely a strong connection between shamanism and ritual-based influence with both prestige and dominance styles of leadership. Ritual and supernatural capacities may represent a dimension of leader qualities in which dominance and prestige are coupled forming an alternative style of leadership, as shamans often provide a necessary and valued public good, have special expertise and abilities, are feared and achieve or maintain influence via threats or use of direct and/or indirect aggression. Artistic performance is also implicated in this style of leadership and ceremonial, artistic performances by shamans are common displays which often incorporate the superhuman abilities of shamans [@singh_cultural_2017].

@@

<!-- DISCUSSES SUCCESS FEATURE -->
**Leader success**. Leaders, probably as a consequence of their abilities to provide social benefits (e.g., resolving conflicts and being generous) and/or impose social costs (e.g., punishing and aggressiveness), are also characterized by their material, social, and reproductive success (bottom cluster Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature focused on associated outcomes of leadership and social hierarchy [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013]. Several of these measures of success often co-occur. Among the Trobriands, horticulturalists in Papua New Guinea, @ol06-037 [p. 47] describes relationships between polygynous marriages, increased wealth, and widespread social contacts:

<!-- DISCUSSES BOTH MANAGEMENT FEATURES -->
**Leaders are managers**. In the absence of state-institutions local leaders are often tasked with managing various economic systems and providing social services. Within the domain of leader management, there are likely to be two primary features of human leader functions: strategizing and organizing. Our results correspond very closely to what @van_vugt_cognitive_2007 define as *strategic* leadership, where leaders increase the pay-offs of another behavior thereby incentivizing followership versus *coordinating* leadership, where individuals bound by group challenges must effectively organize collective effort.

@@

<!-- PROSCIAL FUNCTIONS FEATURE -->
**Public goods**. Beyond strategically shaping inter-individual behaviors or organizing collective behavior, we found much of leadership involves beneficial investments in group welfare. The prosocial feature of leader functions included distributing resources, providing protection, and various prosocial investments, which are functions commonly highlighted in the anthropological literature on leadership in nonindustrial societies [@sahlins_poor_1963; @johnson_organizational_1982; @service_origins_1975; @boehm_egalitarian_1993; @boehm_hierarchy_1999; @johnson_evolution_1987]. In the absence of institutionalized leadership roles, which characterize many of the relatively small-scale populations in this sample, leaders often maintain social influence via continued community investment and the demands of followers can shape the functions of leaders to meet local needs.

<!-- MAYBE IN CONCLUSION -->
Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, in order to organize group members, implement strategic actions, and provide prosocial services to the group, all while conforming to cultural norms. Dichotomizing this complexity into two distinct social strategies, although theoretically and empirically convenient, likely fails to capture much of the ethnographic reality of human leadership dynamics.

@@

[NEED TO INTEGRATE]
These results provide a framework for further work on evolutionarily relevant costs and benefits of leadership, and suggests leaders-follower dynamics should continue to be prioritized in evolutionary models of demography, group dynamics, social complexity and inequality, cooperation and punishment, and resource exchange [c.f., @bird_variability_2019; @ember_our_2018; @micheletti_modelling_2020; @smith_cultural_2020; @birch2017philosophy; @marlowe_more_2008; @henrich_markets_2010; @mace_co-evolution_1998; @mace_phylogenetic_2005; @mattison_evolution_2016]

@@

# Concluding remarks: Beyond the non-WEIRD 
<!-- NOT SURE OF TITLE -- WANT TO EMPHASIZE NON WEIRD DIVERSITY AND ALSO MENTION BACK TO WERID FINDINGS BRIEFLY -->
In recognizing the limitations of most extant psychological, sociological, and managerial data, many social scientists are now diversifying their samples across nations, cultures, and contexts to better understand *human* psychology, behavior, and culture. We caution that the WEIRD people problem in behavioral science cannot be remedied simply by a non-WEIRD solution. As our results, as well as much of anthropology and ethnography, demonstrate, within the populations that are now often coined a non-WEIRD, there is enormous diversity. Furthermore, larger order categories such as "subsistence strategy" are probably less relevant for understanding behavioral diversity than more salient group contexts. Despite limitations, the ethnographic record present an invaluable opportunity to explore patterns of human diversity and evolutionary theorists and researchers looking beyond WEIRD samples should take special interest in such methods.

We found leaders in the ethnographic record are portrayed as generally prosocial, serve to resolve conflicts and organize cooperation, and receive material, social, and reproductive benefits. These results conform to the characteristics of leaders that are emphasized in the literature on the evolution of leadership and provide broad cross-cultural support for several influential theories, such as the prestige-dominance model, the collective action model, the service-for-prestige model, and the computational services model [for review, see @garfield_evolutionary_2019].

While there is a evidence of a "dominant" style of leadership involving coercive authority, confidence and aggression, as well as a "prestige" style involving special knowledge, experience, and innovation, we suggest rethinking models of leadership styles and focusing on inter-individual variation in specific competencies. There is likely to be substantial overlap in physical, spiritual, cognitive, and other prosocial endowments and the degree to which a leader relies on any will be shaped by their individual levels of such capacities, the presence and capacities of their rivals, as well as group context and cultural pressures. For example, widespread preferences for egalitarianism and social equality are likely strong drivers of leader preferences across diverse societies [@cheng_dominance_2019], but shamanism and ritual based contexts represent a social domain in which ideologies support a style of leadership fusing both dominance and prestige-based competencies.

<!-- One interesting pattern that has not been widely discussed in the evolutionary literature on leadership is the role of shamans and their intersection, in our results, with the dominant style of leadership. Our social costs cluster (Figure \@ref(fig:cluster)A) includes the *Supernatural* abilities variable. @winkelman1987cross found that in 47% of societies sampled from the HRAF ($n = 43$) shamans were associated with providing decision-making services and associated with military power. Providing ritual functions and military leadership are common (documented in over 50% of societies here) functions leaders serve (see Figure \@ref(fig:varsupport)). -->

<!-- We also found a link between prestige-based leadership, which is associated with deference towards accomplished individuals with expertise, special skills, and high intelligence [@cheng_dominance_2019; @henrich_evolution_2001], and charismatic leadership, which is associated with oratory skills, charismatic personality, confidence, and interpersonal skills [@grabo_charismatic_2016; @howell_role_2005]. See, Figure \@ref(fig:cluster)A, *Ability to provide social benefits* cluster. This suggests that these leadership strategies might be closely related across a broad range of societies and social contexts, or part of a broader strategy of social influence based on the provisioning of various benefits. -->

<!-- THINK WE ALREADY EXPRESSED THIS BELOW -->
The strong support for material and social benefits associated with leadership (Figure \@ref(fig:varsupportcb)) and the colinearity of measures associated with material wealth, social capital, and reproductive success (Figure \@ref(fig:cluster)A), suggests processes of leadership and followership might be strong drivers of the emergence and intergenerational transmission of inequality [@bowles_emergence_2010; @mattison_evolution_2016]. Various sources of material wealth and strategic or institutionalized processes of nepotism are widespread cross-culturally, including among many societies often described as "egalitarian." These results suggest anthropologists and economists investigating the emergence and evolution of inequality, including material, social, and reproductive dimensions, should incorporate and pay close attention to group processes and social strategies associated with leadership and followership.

@@

The 60-culture Probability Sample Files of the HRAF developed by @naroll_proposed_1967 is based on culturally and geographically diverse populations which are well covered in the ethnographic record. The distribution of cultures, however, is not equal across categories of subsistence strategies; agriculturalists and horticulturalists are relatively over-represented compared to hunter-gatherers and pastoralists. The ethnographic sample of text developed in the *leadershipdata* package is also not equally balanced across all subsistence strategy categories. See Figure S\@ref(fig:subsisttree).

@@

The distribution of cultures by subsistence strategy across continental region is also variable. In this sample, most African cultures are agriculturalists and most North American cultures are hunter-gatherers, for example. See Figure S\@ref(fig:regionsubsistence).
